### 2025-05-07 paper/paper.tex
Update paper.tex

One potential concern over the implementation of our model is not the model itself, but the trustworthiness of the results. As a result, the design of the model took careful consideration into preparing valid results through cross validation and critical analysis. Various situations were considered for training, while only a single traffic scenario was used for testing and benhcmarking.

The first way we validated our results is by creating new scenarios. Each Episode of the model sees a completely new traffic scenario within Starkville, MS, granted with similar size due to computational limitations of large scale traffic. We then printed the reward, the punishment part of the Q-score each episode the ensure progress and exploration, and the results reveal relative consistency with a strong balance of exploration and exploitation. Even with new situations, the score grew and grew over time (please note that the reward is negative). The model was trained for 150 episodes in this scenario, to reach convergance while exploring as much as possible.

@@

During our benchmarking phase we generated one testing scenario. This scenario uses a moderate-light traffic, based on Starkville, Mississippi's population and is used as a baseline to ensure that the improved results are not a result of a more or less congested traffic pattern. When we run the benchmarks we carefully ensured that this was performed to make sure results are trustworthy, and the the model did not opimize only one single situation it saw. When doing this, the results were overwhelmingly positive, and ensures that MARL based Deep Q-Learning is a valid way to optimize traffic signal controllers, as seen in section ``Results over baseline.``

@@

In order to download the code one would need to go to github, and view chandlerweeks/TrafficSignalManagement. The link to the repository is: https://github.com/ChandlerWeeks/TrafficSignalManagement. Once here, fork the repository and make a local clone on your device. If a release is available at the time of reading the paper, that link will allow direct access of the codebase.

The model was not designed within a virtual environment, due to conflicts between operating systems of installations, so one will need to install the dependencies for the code beforehand, most noteably traci, likely via the pip install traci through a terminal.

In order to run the code, SUMO will also need to be installed, following SUMO's tutorial for installation for the necessary operating system. Once installed, the path variables will need to be set for sumo and sumo-gui in order for the code to run.

@@

Once every prerequisite has been setup, to test installation and configuration, one should open SUMO-gui and open the file in data/simulation.sumocfg. If the file runs and traffic is observed, then the testing traffic flow scenario has been loaded and the code is ready to run.

In order to run training, navigate to the TrafficSignalManagement file in a terminal instance and run the provided command. \texttt{\texttt{python traffic\_signal\_nn/main.py --config traffic\_signal\_nn/config/config\_dqn\_city.ini train}}. Common errors include not having SUMO variables added to the path, and not having pip modules installed. Each episode will run, taking about a minute to simulate an hour of traffic. The reward will be printed to the console to let the user observe the loss, to see when the model explores and when it converges. Parameters can be modified for training within traffic\_signal\_nn/config/config\_dqn\_city.ini, including the neural network parameters and the selected roads. The code can take hours at a time to run, and a demo can be provided if necessary, via youtube.

In order to benchmark the baseline for SUMO, the scripts benchmark.py and get\_queue\_length.py were developed. benchmark simply gets the results of the runtime through the file data/output/tripinfo.xml. It contains information about the trip, such as average trip time, average waiting time, and average delay, but does not contain information on average queue length. Before running benchmarks, it is important to run the get queue length script, which obtains the average queue lengths seperate from what SUMO provides. One must run the benchmark scenario provided through simulation.sumocfg within SUMO-gui to get the correct output. In order to run the benchmarks, simply run \texttt{python benchmark.py} after running the simulation, along with \texttt{python .\textbackslash get\_queue\_length.py --config .\textbackslash data\textbackslash simulation.sumocfg --output .\textbackslash data\textbackslash output\textbackslash queue\_length.csv} to determine the queue lengths.

@@

Large language models were used in assistance of coding, particularly for debugging and assistance with small code snippets, but never used to generate large parts of a funciton, or codebits. The code is developed by human, but AI assisted with some lines.

>>>>

One potential concern over the implementation of our model is not the model itself, but the trustworthiness of the results. As a result, the design of the model took careful consideration into preparing valid results through cross validation and critical analysis. Various situations were considered for training, while only a single traffic scenario was used for testing and benchmarking.

The first way we validated our results is by creating new scenarios. Each Episode of the model sees a completely new traffic scenario within Starkville, MS, granted with similar size due to computational limitations of large scale traffic. We then printed the reward, the punishment part of the Q-score each episode the ensure progress and exploration, and the results reveal relative consistency with a strong balance of exploration and exploitation. Even with new situations, the score grew and grew over time (please note that the reward is negative). The model was trained for 150 episodes in this scenario, to reach convergence while exploring as much as possible.

@@

During our benchmarking phase we generated one testing scenario. This scenario uses a moderate-light traffic, based on Starkville, Mississippi's population and is used as a baseline to ensure that the improved results are not a result of a more or less congested traffic pattern. When we run the benchmarks we carefully ensured that this was performed to make sure results are trustworthy, and the the model did not optimize only one single situation it saw. When doing this, the results were overwhelmingly positive, and ensures that MARL based Deep Q-Learning is a valid way to optimize traffic signal controllers, as seen in section ``Results over baseline.``

@@

In order to download the code one would need to go to GitHub, and view chandlerweeks/TrafficSignalManagement. The link to the repository is: https://github.com/ChandlerWeeks/TrafficSignalManagement. Once here, fork the repository and make a local clone on your device. If a release is available at the time of reading the paper, that link will allow direct access of the codebase.

The model was not designed within a virtual environment, due to conflicts between operating systems of installations, so one will need to install the dependencies for the code beforehand, most notably traci, likely via the pip install traci through a terminal.

In order to run the code, SUMO will also need to be installed, following SUMO's tutorial for installation for the necessary operating system. Once installed, the path variables will need to be set for sumo and sumo-GUI in order for the code to run.

@@

Once every prerequisite has been setup, to test installation and configuration, one should open SUMO-GUI and open the file in data/simulation.sumocfg. If the file runs and traffic is observed, then the testing traffic flow scenario has been loaded and the code is ready to run.

In order to run training, navigate to the TrafficSignalManagement file in a terminal instance and run the provided command. \texttt{\texttt{python traffic\_signal\_nn/main.py --config traffic\_signal\_nn/config/config\_dqn\_city.ini train}}. Common errors include not having SUMO variables added to the path, and not having pip modules installed. Each episode will run, taking about a minute to simulate an hour of traffic. The reward will be printed to the console to let the user observe the loss, to see when the model explores and when it converges. Parameters can be modified for training within traffic\_signal\_nn/config/config\_dqn\_city.ini, including the neural network parameters and the selected roads. The code can take hours at a time to run, and a demo can be provided if necessary, via YouTube.

In order to benchmark the baseline for SUMO, the scripts benchmark.py and get\_queue\_length.py were developed. benchmark simply gets the results of the runtime through the file data/output/tripinfo.xml. It contains information about the trip, such as average trip time, average waiting time, and average delay, but does not contain information on average queue length. Before running benchmarks, it is important to run the get queue length script, which obtains the average queue lengths separate from what SUMO provides. One must run the benchmark scenario provided through simulation.sumocfg within SUMO-GUI to get the correct output. In order to run the benchmarks, simply run \texttt{python benchmark.py} after running the simulation, along with \texttt{python .\textbackslash get\_queue\_length.py --config .\textbackslash data\textbackslash simulation.sumocfg --output .\textbackslash data\textbackslash output\textbackslash queue\_length.csv} to determine the queue lengths.

@@

Large language models were used in assistance of coding, particularly for debugging and assistance with small code snippets, but never used to generate large parts of a function, or codebits. The code is developed by human, but AI assisted with some lines.
